Distortion corrected kernel density estimate on Riemannian manifolds

### 2022-11-06 paper/paper.Rmd
Update PDF, table electricityoutliers hash for color codes to be solved

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = T,
  messages = FALSE,
  warning = FALSE,
  eval = TRUE
  # include = FALSE
)
options(tinytex.verbose = TRUE)
library(tidyverse)
library(here)
# library(hdrcde)
# library(car)
library(patchwork)
library(kableExtra)
# Jmisc::sourceAll(here::here("R/sources"))
# set.seed(1234)
```

@@

```{r electricityembedding, fig.align = 'center', out.width = "90%", fig.cap="Highest density region plots of the 2-D embedding for $3,639$ households, with each point representing the distribution of one household and colors indicating the density estimated from DC-KDE (left panel) and KDE (right panel). The black points with blue text indexing are the top 20 anomalous households found by each dimension reduction method in each row.", echo=FALSE, eval=T}
knitr::include_graphics("figures/Electricity_2d_5levels_outliers_comparison_5ml_r180.png")
```
```{r electricityoutliers}
load("figures/Electricity_2d_hdrtable_umapvs4ml_radius20_r180.rda")
tabletop20
```

@@

```{r electricityhdrpercent}
load("figures/Electricity_2d_hdrtable_umapvs4ml_radius20_r180.rda")
table_outlierpercent
```
```{r electricityrank, fig.align = 'center', out.width = "90%", fig.cap="Highest density region plots of the 2-D embedding for $3,639$ households, with each point representing the distribution of one household and colors indicating the density estimated from DC-KDE (left panel) and KDE (right panel). The black points with blue text indexing are the top 20 anomalous households found by each dimension reduction method in each row.", echo=FALSE, eval=T}
knitr::include_graphics("figures/Electricity_2d_rankplot_3639id_umapvs4ml_radius20_r180.png")
```

@@

```{r electricityoutliers, fig.align = 'center', out.width = "90%", fig.cap="Electricity usage plots of all 535 days for the most typical household and two anomalies in rows and KDE with distortion correction (left panel) and KDE (right panel) in columns.", echo = FALSE, eval=T}
knitr::include_graphics("figures/Electricity_2d_compare4id_1typical_3anomalous.png")
```

Now let us take a close look at the typical and anomalous distributions of households. According to \autore{tab:electricityoutliers}, we pick the three dark blue indexes that are identified as anomalies the most times, which are Household 3161, 4546, 7049, as well as the most typical household with the highest density estimate, Households 2396, for illustration. Their half-hourly electricity demands across the entire trial period are plotted in \autoref{fig:electricityoutliers}. Compared with ID 2396, the other three households show different usage patterns. ID 3161 has much less electricity usage in summers, mostly below 1kWh, than in winters, above 3kWh. As this data set does not include the energy used for cooling or heating systems, it is possible that there are other large electronic appliances used in winter for this household. The other two anomalous IDs, 4546 and 7049, however, have nearly zero kWh electricity consumption for many days and some spikes for other days. It is likely that these two households are not occupied all the time during the trial and could be used as a AirBnb place or holiday place.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = T,
  messages = FALSE,
  warning = FALSE,
  eval = TRUE,
  include = TRUE
)
options(tinytex.verbose = TRUE)
library(tidyverse)
library(here)
library(patchwork)
library(kableExtra)
# Jmisc::sourceAll(here::here("R/sources"))
# set.seed(1234)
```

@@

```{r electricityembedding, fig.align = 'center', out.width = "90%", fig.cap="Highest density region plots of the 2-D embedding for $3,639$ households, with each point representing the distribution of one household and colors indicating the density estimated from DC-KDE (left panel) and KDE (right panel). The black points with blue text indexing are the top 20 anomalous households found by each dimension reduction method in each row.", echo=FALSE, eval=T}
knitr::include_graphics("figures/Electricity_2d_5levels_outliers_comparison_5ml_r180.png")
```
```{r electricityoutliers, eval=F}
load("figures/Electricity_2d_table_outlier_idcompare_5ml_radius20_r180.rda")
# f20cols <- paste0("\\", f20cols)
# tabletop20 <- 
  f20 %>% 
  kableExtra::kbl(caption = "Top 20 anomalous household IDs using density estimates of different 2-dimensional embedding. The left five columns are anomalies by DC-KDE, while the right five columns are anomalies using KDE. The blue colored IDs are IDs that are detected over three times, with the darker one indicating five times and the lighter one indicating four times.", booktabs = TRUE, digits = 3, escape = FALSE, col.names = NULL, align=rep('c', 10)) %>%
  # kable_styling(latex_options = "scale_down") %>%
  kable_paper(full_width = TRUE) %>%
  add_header_above(c(" " = 1, rep(c("ISOMAP" = 1, "LLE" = 1, "Laplacian Eigenmaps" = 1, "t-SNE" = 1, "UMAP" = 1), 2))) %>% 
  add_header_above(c(" " = 1, "DC-KDE" = 5, "KDE" = 5)) %>%
  # row_spec(1, bold = TRUE) %>% 
  column_spec(1, bold = FALSE) %>% 
  column_spec(2, bold = freqmax[,2], color = f20cols[,2]) %>% 
  column_spec(3, bold = freqmax[,3], color = f20cols[,3]) %>% 
  column_spec(4, bold = freqmax[,4], color = f20cols[,4]) %>% 
  column_spec(5, bold = freqmax[,5], color = f20cols[,5]) %>% 
  column_spec(6, bold = freqmax[,6], color = f20cols[,6]) %>% 
  column_spec(7, bold = freqmax[,7], color = f20cols[,7], border_left = TRUE) %>% 
  column_spec(8, bold = freqmax[,8], color = f20cols[,8]) %>% # background = grey(0.9)
  column_spec(9, bold = freqmax[,9], color = f20cols[,9]) %>% 
  column_spec(10, bold = freqmax[,10], color = f20cols[,10]) %>% 
  column_spec(11, bold = freqmax[,11], color = f20cols[,11])
# tabletop20
```

@@

```{r electricityhdrpercent}
load("figures/Electricity_2d_hdrtable_umapvs4ml_radius20_r180.rda")
# table_outlierpercent <- 
hdrtable %>% 
  head(n = 2) %>% 
  `rownames<-` (c(">99\\% HDR", "99\\% HDR")) %>% 
  kableExtra::kbl(caption = "Proportion of the same outliers as UMAP in two highest density regions when estimating densities for four manifold learning embeddings.", booktabs = TRUE, digits = 3, escape = FALSE, align = "c",
                  # col.names = c(rep(c("DC-KDE", "KDE"), 4))
  ) %>%
  # kable_classic("striped") %>%
  kable_styling(latex_options = "scale_down") %>%
  kable_paper(full_width = TRUE) %>%
  add_header_above(c("HDRs" = 1, rep(c("DC-KDE" = 1, "KDE" = 1), 4))) %>%
  add_header_above(c(" " = 1, "ISOMAP" = 2, "LLE" = 2, "Laplacian Eigenmaps" = 2, "t-SNE" = 2)) %>% 
  # column_spec(1, bold = TRUE) %>% 
  # add_indent(1:6) %>% 
  column_spec(2, bold = hdrmax[,1]) %>% 
  column_spec(3, bold = hdrmax[,2]) %>% 
  column_spec(4, bold = hdrmax[,3]) %>% 
  column_spec(5, bold = hdrmax[,4]) %>% 
  column_spec(6, bold = hdrmax[,5]) %>% 
  column_spec(7, bold = hdrmax[,6]) %>% 
  column_spec(8, bold = hdrmax[,7]) %>% 
  column_spec(2*(1:4), border_left = T, background = grey(0.9))
# table_outlierpercent
```
```{r electricityrank, fig.align = 'center', out.width = "90%", fig.cap="Highest density region plots of the 2-D embedding for $3,639$ households, with each point representing the distribution of one household and colors indicating the density estimated from DC-KDE (left panel) and KDE (right panel). The black points with blue text indexing are the top 20 anomalous households found by each dimension reduction method in each row.", echo=FALSE, eval=T}
knitr::include_graphics("figures/Electricity_2d_rankplot_3639id_umapvs4ml_radius20_r180.png")
```

@@

```{r electricityraw, fig.align = 'center', out.width = "90%", fig.cap="Electricity usage plots of all 535 days for the most typical household and two anomalies in rows and KDE with distortion correction (left panel) and KDE (right panel) in columns.", echo = FALSE, eval=T}
knitr::include_graphics("figures/Electricity_2d_compare4id_1typical_3anomalous.png")
```

Now let us take a close look at the typical and anomalous distributions of households. According to \autoref{tab:electricityoutliers}, we pick the three dark blue indexes that are identified as anomalies the most times, which are Household 3161, 4546, 7049, as well as the most typical household with the highest density estimate, Households 2396, for illustration. Their half-hourly electricity demands across the entire trial period are plotted in \autoref{fig:electricityraw}. Compared with ID 2396, the other three households show different usage patterns. ID 3161 has much less electricity usage in summers, mostly below 1kWh, than in winters, above 3kWh. As this data set does not include the energy used for cooling or heating systems, it is possible that there are other large electronic appliances used in winter for this household. The other two anomalous IDs, 4546 and 7049, however, have nearly zero kWh electricity consumption for many days and some spikes for other days. It is likely that these two households are not occupied all the time during the trial and could be used as a AirBnb place or holiday place.
